An algorithm for searching the global minimum of univariate discontinuous functions\thanks{This work was supported by project no. 0729-2020-0055.}

### 2021-06-10 2021/ICCS_BarkalovUsova/iccs.tex
Include one more comment in Russian

\usepackage{hyperref}

@@

%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
%\usepackage[utf8x]{inputenc}
%\usepackage[english,russian]{babel}
%\usepackage{cmap}
%%%%

@@

This paper considers the problem of finding the global minimum $x^*$ of a one-dimensional function $\varphi(x)$ of the form 
\begin{equation}\label{problem}
\varphi(x^*)=\min\left\{\varphi(x):x\in\left[a,b\right]\right\}.
\end{equation}
One-dimensional optimization problems play an important role, because many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).

The literature describes various methods for solving the problem (\ref{problem}) depending on the assumptions about the properties of the objective function \cite{Horst1995,Horst1996,Pinter1996}. One of such assumptions, which is often fulfilled in applied optimization problems, is that the objective function $\varphi(x)$ satisfies the Lipschitz condition 
\[
\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|,\; x',x'' \in [a,b],
\]
where the constant $L$ is unknown a priori.

This feature of the objective function is commonly used in many approaches to the development of algorithms for multiextremal optimization \cite{Evtushenko2009,Evtushenko2013,Jones2009,Paulavicius2011}.

>>>>

%\usepackage{hyperref}

@@

%%%% ДЛЯ РУССКОГО ТЕКСТА закомментировать потом!
\usepackage{inputenc}
\usepackage[english,russian]{babel}
%\usepackage{cmap}
%%%%

@@

This paper considers the problem of finding the global minimum $x^*$ of a one-dimensional function $\varphi(x)$ of the form 
\begin{equation}\label{problem}
\varphi(x^*)=\min\left\{\varphi(x):x\in\left[a,b\right]\right\}.
\end{equation}

Задачи глобальной оптмизации, которые характеризуются наличием многих локальных экстремумов, возникают во многих областях приложений (see, e.g., \cite{Cavoretto2021,Kvasov2015,Kvasov2013,Modorskii2016}). Как правило, вычисление даже одного значения целевой функции в таких задачах характеризуется значительным временем, т.к. требует трудоемкого численного моделирования. В этом случае принято говорить о expensive black-box optimization problems. Задачи указанного класса чрезвычайно сложны и являются объектом интенсивных исследований.

Следует отметить важную роль одномерных задач глобальной оптимизации. Во-первых, как уже отмечалось, они возникают во многих приложениях (see, e.g., \cite{Gillard2017,Sergeyev2020}). А во-вторых,  
 many approaches to solving multidimensional optimization problems are, in one way or another, based on reducing the solution of the original multidimensional problem to solving a series of related one-dimensional optimization problems (see, for example, popular approaches in \cite{Pinter1996,PaulaviciusZilinskas2014,Sergeyev2017,Sergeyev2013}).

The literature describes various methods for solving the problem (\ref{problem}) depending on the assumptions about the properties of the objective function \cite{Horst1995,Horst1996,Pinter1996}. One of such assumptions, which is often fulfilled in applied optimization problems, is that the objective function $\varphi(x)$ satisfies the Lipschitz condition 
\[
\left|\varphi(x')-\varphi(x'')\right|\leq L\left|x'-x''\right|,\; x',x'' \in [a,b],
\]
where the constant $L$ is unknown a priori. 
This feature of the objective function is commonly used in many approaches to the development of algorithms for multiextremal optimization \cite{Evtushenko2009,Evtushenko2013,Jones2009,Paulavicius2011}.
